Schema Change: Challenge Problems

### 2024-05-01 paper.tex
use biblatex, comments
natbib isn't working. Journal template uses biblatex so switched to that. still an error about inputenx

\documentclass[english,submission]{programming}
%\usepackage[backend=biber]{biblatex}
%\addbibresource{paper.bib}

\usepackage{changepage}
\usepackage{multicol}
\usepackage{ccicons}
\usepackage{natbib}
\usepackage[many]{tcolorbox}

@@

@@

\author[e]{Geoffrey Litt}%[0000-0003-0858-5165]
\authorinfo{is TBD. Contact him at \email{gklitt@gmail.com}.}
\affiliation[b]{Ink \& Switch, Planet Earth}

@@

\keywords{schema change, live programming, local-first programming}

@@

\noindent\textbf{Context}: \emph{What is the broad context of the work? What is the importance of the general research area?}
Schema change is unsolved problem in live and local-first programming. It is also poorly understood
problem that calls for a more fundamental analysis. It is particularly important as more programming
is done in programming systems where programs may be changed by their users and as more programming
is local-first, meaning that such changes need to be synchronized.

@@

Schema change won't go away. Changing requirements and changing code lead to changing the schema of a database. Schema change, also called schema migration, is the problem of migrating existing data from the old schema to the new. This often involves custom migration programs or specialized Domain Specific Langiuages. The migration must be carefully coordinated with upgrading the application code and associated artifacts that assume the new schema. Schema change on database servers is often delegated to Database Administrators and DevOps. Live programming~\cite{tanimoto90,hancock03} and local-first software~\cite{localfirst} both move data away from the server, eliminating those jobs, but not eliminating the need to do schema change, and indeed increasing the need to do it automatically. Recently \citet{Cambria} spotlighted these problems and proposed an approach using lenses~\cite{Foster2007}, but otherwise there has been surprisingly little research. To promote further progress we offer a set of challenge problems to the live programming and local-first communities, inviting them to propose and compare solutions.

@@

@@

\textcolor{red}{TODO: Maybe the right way to frame this is to say that the challenges are intended
less for evaluating particular solutions but more for framing and making sense of different
kinds of schema-chaange-related problems.  You can read this paper and when you face a schema-change-related
problem, think about it in terms of our dimensions. Maybe compare it to our challenges.
We do not necessarilly expect that people will directly solve our challenges though. (Although
some might and that would be nice..)}

@@

\begin{figure}[t]
\vspace{-1em}
\centering
\includegraphics[width=0.5\textwidth]{figures/layers.png}
\caption{Pace layers from Stewart Brand's classic \emph{How Buildings Learn} \cite{Brand95}.
  The outer layers rarely change, while the inner layers change frequently. The geographic
  site of a building is eternal, its structure remains stable for decades, space plan
  needs to change every couple of years. ``Because of the different rates of change of its
  components, a building is always tearing itself apart.''}
\label{fig:layers}
\end{figure}

@@

\bibliographystyle{abbrv}
\bibliography{paper}

>>>>

\documentclass[english,submission]{programming}

\usepackage{csquotes}
\usepackage{inputenx}
\usepackage[backend=biber]{biblatex}
%\usepackage{natbib}
\addbibresource{paper.bib}

\usepackage{changepage}
\usepackage{multicol}
\usepackage{ccicons}
\usepackage[many]{tcolorbox}

@@

% define natbib citet
\newcommand{\citet}[1]{\citeauthor*{#1} \cite{#1}}

@@

\author[e]{Geoffrey Litt}%[0000-0003-0858-5165]
\authorinfo{is TBD. Contact him at \email{gklitt@gmail.com}.}
\affiliation[e]{Ink \& Switch, Planet Earth}

@@

\keywords{schema change, database migration, live programming, local-first programming}

@@

\noindent\textbf{Context}: \emph{What is the broad context of the work? What is the importance of the general research area?}
Schema change is an unsolved problem in live and local-first programming that is poorly understood and calls for a more fundamental analysis. It is particularly important as more programming
is done in programming systems where programs may be changed by their users and as more programming
is local-first, meaning that such changes need to be synchronized.

@@

Schema change won't go away. Changing requirements and changing code lead to changing the schema of a database. Schema change, also called schema migration, is the problem of migrating existing data from the old schema to the new. This often involves custom migration programs or specialized Domain Specific Langiuages. The migration must be carefully coordinated with upgrading the application code and associated artifacts that assume the new schema. Schema change on database servers is often delegated to Database Administrators and DevOps. Live programming~\cite{tanimoto90,Hancock03} and local-first software~\cite{localfirst} both move data away from the server, eliminating those jobs, but not eliminating the need to do schema change, and indeed increasing the need to do it automatically. Recently \citet{Cambria} spotlighted these problems and proposed an approach using lenses~\cite{Foster2007}, but otherwise there has been surprisingly little research. To promote further progress we offer a set of challenge problems to the live programming and local-first communities, inviting them to propose and compare solutions.

@@

~
\textcolor{red}{[JE] We could expand our context to include conventional DB systems not just live/local. It is a widespread problem, attested by the many tools to assiste with it. We could include in our problems a description of a typical solution with a standard DB as a sort of baseline. }

@@

\textcolor{red}{TODO: Maybe the right way to frame this is to say that the challenges are intended
less for evaluating particular solutions but more for framing and making sense of different
kinds of schema-change-related problems.  You can read this paper and when you face a schema-change-related
problem, think about it in terms of our dimensions. Maybe compare it to our challenges.
We do not necessarilly expect that people will directly solve our challenges though. (Although
some might and that would be nice..)}

@@

\begin{figure}[t]
\vspace{-1em}
\centering
\includegraphics[width=0.5\textwidth]{figures/layers.png}
\caption{Pace layers from Stewart Brand's classic \emph{How Buildings Learn} \cite{Brand95}.
  The outer layers rarely change, while the inner layers change frequently. The geographic
  site of a building is eternal, its structure remains stable for decades, space plan
  needs to change every couple of years. ``Because of the different rates of change of its
  components, a building is always tearing itself apart.''
  \textcolor{red}{[JE] Except the outer skin changes faster than the structure.}
  }
\label{fig:layers}
\end{figure}

\textcolor{red}{[JE] `change in a program' might be misunderstood as changes to the source code. Perhaps `system' or `software system' might be better than `program'? }

@@

%\bibliographystyle{abbrv}
%\bibliography{paper}
\printbibliography
